---
title: "MIKK panel fecundity"
author: "Ian Brettell"
date: '`r format(Sys.Date())`'
output:
  html_document:
    toc: true
    toc_float: true
    dev: 'svg'
    number_sections: true
    keep_md: true
    pandoc_args: --lua-filter=color-text.lua
    highlight: pygments  
---

# Setup

```{r, message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(plotly)
library(scales)
library(lubridate)
library(cowplot)
```

# Quantitative data

## Read in data

```{r}
in_file = here::here("data", "fecundity", "20210205_quantitative.xlsx")
df_quant = readxl::read_xlsx(in_file)
```

## Tidy data

```{r, results = 'asis'}
# File to save to 
out_file = here::here("data", "fecundity", "20210208_quant.csv")

# Remove NAs
df_quant = df_quant %>% 
  filter(!is.na(Strain))

# Split into list
quant_list = list("F15" = df_quant[, 1:8],
                  "F16" = df_quant[, c(1, 9:length(df_quant))])

# Pivot longer
out_df = lapply(quant_list, function(generation){
  out = generation %>% 
    # Pivot longer
    tidyr::pivot_longer(cols = contains("eggs"),
                        names_to = "DATE",
                        names_prefix = "eggs ",
                        values_to = "EGG_COUNT") %>% 
    # Convert into date
    dplyr::mutate(DATE = DATE %>% 
                  str_c("2020") %>% 
                  str_replace_all("\\.", "-") %>% 
                  as.Date("%d-%m-%Y")) %>% 
    # Get weekday
    dplyr::mutate(WEEKDAY = weekdays(DATE)) %>% 
    # Rename columns
    dplyr::select(STRAIN = "Strain",
                  FEMALE_COUNT = contains("females"),
                  DATE, WEEKDAY, EGG_COUNT,
                  everything()) %>% 
    # Replace question marks in FEMALE_COUNT and convert to integer
    dplyr::mutate(FEMALE_COUNT = str_replace(FEMALE_COUNT, "\\?", "") %>% 
                    as.integer()) %>% 
    # Get eggs per female
    dplyr::mutate(EGGS_PER_FEMALE = EGG_COUNT / FEMALE_COUNT)
  
  return(out)
}) %>% 
  # bind in to DF
  dplyr::bind_rows(.id = "GENERATION") %>% 
  # Write to CSV
  readr::write_csv(out_file, na = "")

# Adapt variables for plotting
strain_levels = unique(out_df$STRAIN)
pal_primary = hue_pal()(length(strain_levels))
names(pal_primary) = strain_levels

out_df$STRAIN = factor(out_df$STRAIN, levels = strain_levels)
out_df$WEEKDAY = factor(out_df$WEEKDAY,
                        levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

knitr::kable(head(out_df))
```

## Plot

### Correlation between F15 and F16

```{r}
corr_plot = out_df %>% 
  dplyr::group_by(GENERATION, STRAIN) %>% 
  summarise(mean(EGGS_PER_FEMALE)) %>% 
  tidyr::pivot_wider(names_from = GENERATION,
                     values_from = `mean(EGGS_PER_FEMALE)`) %>% 
    ggplot() +
      geom_point(aes(F15, F16, colour = STRAIN)) +
      coord_fixed() +
      theme_bw() +
      guides(colour = F) +
      xlim(0,7) +
      ylim(0,7) +
      ggtitle("Mean eggs per female") +
      scale_color_manual(values = pal_primary)

# Plotly
ggplotly(corr_plot, height = 400, width = 400) %>% 
  layout(showlegend = F)
```

### Effect of day of collection

```{r}
collection_day_plot = out_df %>% 
  ggplot() +
    geom_point(aes(DATE, EGGS_PER_FEMALE, colour = STRAIN), alpha = 0.8) +
    facet_wrap(vars(GENERATION, WEEKDAY)) +
    guides(colour = F) +
    theme_bw() +
    xlab("Date") +
    ylab("Mean eggs produced per female") +
    scale_color_manual(values = pal_primary)

ggplotly(collection_day_plot, height = 1000, width = 800) %>% 
  layout(showlegend = F)
```


```{r}
# Get function to reverse labels in ggplotly (from https://stackoverflow.com/questions/59611914/reverse-the-legend-order-when-using-ggplotly)
reverse_legend_labels <- function(plotly_plot) {
  n_labels <- length(plotly_plot$x$data)
  plotly_plot$x$data[1:n_labels] <- plotly_plot$x$data[n_labels:1]
  plotly_plot
}

violin_plot = out_df %>%
  dplyr::mutate(STRAIN = factor(STRAIN, levels = rev(strain_levels))) %>% 
  ggplot() +
    geom_violin(aes(STRAIN, EGGS_PER_FEMALE, fill = STRAIN, colour = STRAIN)) +
    geom_jitter(aes(STRAIN, EGGS_PER_FEMALE, label = GENERATION), size = 0.25) +
    theme_bw() +
    coord_flip() +
    guides(fill = F, colour = F) +
    xlab("MIKK panel line") +
    ylab("Mean eggs produced per female") +
    scale_color_manual(values = pal_primary) +
    scale_fill_manual(values = pal_primary)
    

violin_plot %>%
  plotly::ggplotly(height = 1200, width = 700) %>%
  reverse_legend_labels() %>% 
  layout(showlegend = F)
```

# Semi-quantitative data

## Read in data and tidy 

```{r, results = 'asis'}
in_file = here::here("data", "fecundity", "20210205_semiquantitative.xlsx")
df_semi = readxl::read_xlsx(in_file, range = "A1:C81")
out_file = here::here("data", "fecundity", "20210208_semiquant.csv")

# One sample is missing from this dataset. Which one?
df_quant$Strain[which(!df_quant$Strain %in% df_semi$Pair)]

# Create recode vector
date_recode = c("Feb 2019", "Jul 2020")
names(date_recode) = c("2/19", "7/20")
recode_vec_1 = c(0, 1, 2, 3, 4, 5)
names(recode_vec_1) = c(0, "o", "x", "x/", "xx", "xxx")
recode_vec_2 = c("Not producing",
                 "Do not produce every day; <3 eggs when they do",
                 "Do not produce every day; <5 eggs when they do",
                 "0-3 eggs per day",
                 "0-5 eggs per day",
                 "5-10 eggs per day")
names(recode_vec_2) = c(0, 1, 2, 3, 4, 5)
recode_vec_3 = gsub("; ", ";\n", recode_vec_2)

# Tidy

semi_out = df_semi %>% 
  # pivot fecundity
  tidyr::pivot_longer(cols = contains("fecundity"), 
                      names_to = "DATE",
                      names_prefix = "fecundity ",
                      values_to = "FECUNDITY") %>% 
  # recode fecundity measures
  dplyr::mutate(DATE = dplyr::recode(DATE, !!!date_recode),
                FECUNDITY = dplyr::recode(FECUNDITY, "xx/" = "xx"),
                FECUNDITY = dplyr::na_if(FECUNDITY, "do not prod. Yet"),
                FECUNDITY = ifelse(is.na(FECUNDITY), 0, FECUNDITY),
                FECUNDITY = dplyr::recode(FECUNDITY, !!!factor(recode_vec_1)),
                KEY = dplyr::recode(FECUNDITY, !!!recode_vec_2)) %>%
  # rename STRAIN
  dplyr::rename(STRAIN = Pair) %>% 
  # factorise
  dplyr::mutate(STRAIN = factor(STRAIN, levels = strain_levels)) %>% 
  # write to file
  readr::write_csv(out_file, na = "")

knitr::kable(head(semi_out))
```

## Plot

```{r}
semi_all = semi_out %>% 
  dplyr::mutate(STRAIN = factor(STRAIN, levels = rev(strain_levels)),
                KEY = gsub("; ", ";\n", KEY)) %>%
  ggplot() +
    geom_col(aes(STRAIN, KEY, fill = STRAIN)) +
    theme_bw() +
    scale_fill_manual(values = pal_primary) +
    facet_wrap(vars(DATE), ncol = 2) +
    guides(fill = F) +
    coord_flip() +
    theme(axis.text.x = element_text(size = 4)) +
    xlab("MIKK line") +
    ylab("Fecundity")

ggplotly(semi_all, width = 1200, height = 1000) %>% 
  reverse_legend_labels() %>% 
  layout(showlegend = F)
```

### Correlation

```{r}
semi_corr = semi_out %>% 
  tidyr::pivot_wider(id_cols = STRAIN,
                     names_from = DATE,
                     values_from = FECUNDITY) %>% 
  ggplot() +
    geom_jitter(aes(`Feb 2019`, `Jul 2020`, colour = STRAIN), alpha = 0.7) +
    coord_fixed() +
    theme_bw() +
    guides(colour = F) +
    ggtitle("Correlation in semi-quantitative measure") +
    scale_color_manual(values = pal_primary)

  
# Plotly
ggplotly(semi_corr, height = 400, width = 400) %>% 
  layout(showlegend = F)
```

### Final quantitative

```{r}
semi_final = semi_out %>% 
  dplyr::mutate(STRAIN = factor(STRAIN, levels = rev(strain_levels)),
                KEY = gsub("; ", ";\n", KEY)) %>% 
  dplyr::filter(DATE == "Jul 2020") %>% 
  ggplot() +
    geom_col(aes(STRAIN, KEY, fill = STRAIN)) +
    theme_bw() +
    scale_fill_manual(values = pal_primary) +
    guides(fill = F) +
    coord_flip() + 
    theme(axis.text.x = element_text(size = 4)) +
    xlab("MIKK panel line") +
    ylab("Fecundity") +
    ggtitle("Fecundity of the MIKK panel lines as of July 2020")

ggplotly(semi_final, width = 600, height = 1000, tooltip = c("STRAIN", "KEY")) %>% 
  reverse_legend_labels() %>% 
  layout(showlegend = F)
```

Horizontal

```{r}
final_hor = semi_out %>% 
  dplyr::mutate(KEY = gsub("; ", ";\n", KEY),
                KEY = factor(KEY, levels = recode_vec_3)) %>% 
  dplyr::filter(DATE == "Jul 2020") %>% 
  ggplot() +
    geom_col(aes(STRAIN, KEY, fill = STRAIN)) +
    theme_bw() +
    scale_fill_manual(values = pal_primary) +
    guides(fill = F) +
    theme(axis.text.x = element_text(size = 3.5,
                                     angle = 45),
          axis.text.y = element_text(size = 5)) +
    xlab("MIKK panel line") +
    ylab(NULL) +
    ggtitle("Fecundity of the MIKK panel lines as of July 2020")

final_hor
```


```{r}
ggplotly(final_hor, width = 900, height = 300) %>% 
  layout(showlegend = F)
```

# Stacked density plot for inbreeding

Data saved here (copied from Felix's table in the paper):
`data/fecundity/20210212_inbreeding.csv`

```{r}
raw_dat = here::here("data", "fecundity", "20210212_inbreeding.csv")
inbreed_df = readr::read_csv(raw_dat)

total_crosses = 253

inbreed_df_tidy = inbreed_df %>%
  # Remove final row, which makes `N of strains lost` character
  dplyr::slice(-c(11,12)) %>% 
  # Change column names
  dplyr::select(GENERATION = `Inbreeding (Generation)`,
                N_LOST_TOTAL = `N of strains lost`,
                everything()) %>% 
  # Make N_LOST integer
  dplyr::mutate(N_LOST_TOTAL = as.integer(N_LOST_TOTAL)) %>% 
  # Remove MIKK column
  dplyr::select(-MIKK) %>% 
  # Get cumulative lost and N_SURVIVING
  dplyr::mutate(N_LOST_CUM = cumsum(N_LOST_TOTAL),
                N_SURVIVING = total_crosses - N_LOST_CUM) %>% 
  # Replace 
  dplyr::mutate(dplyr::across(c(`not productive`,
                                `infertile male`,
                                `no males`,
                                `no females`,
                                `all cause death`),
                              ~tidyr::replace_na(.x, 0))) %>% 
  # Get cumulative lost for each category
  dplyr::mutate(dplyr::across(c(`not productive`,
                                `infertile male`,
                                `no males`,
                                `no females`,
                                `all cause death`),
                              ~cumsum(.x))) %>% 
  # Amend GENERATION column
  dplyr::mutate(GENERATION = stringr::str_extract(GENERATION, "(?:F\\d{1,2})")) %>% 
  # Add first row
  rbind(., c("F1", rep(0,7), total_crosses)) %>% 
  # Add rows for F12-14
  rbind(., dplyr::filter(., GENERATION == "F11") %>%
            dplyr::slice(rep(1:n(), each = 3)) %>%
            dplyr::mutate(GENERATION = c("F12", "F13", "F14"))) %>% 
  # Factorise GENERATION
  dplyr::mutate(GENERATION = factor(GENERATION, levels = paste("F", seq(1:14), sep = ""))) %>% 
  # Order
  .[order(.$GENERATION), ] %>% 
  # Gather 
  tidyr::pivot_longer(cols = c(`not productive`,
                               `infertile male`,
                               `no males`, 
                               `no females`,
                               `all cause death`,
                               N_SURVIVING),
                      names_to = "STATUS",
                      values_to = "N") %>% 
  # Convert `N` to integer
  dplyr::mutate(N = as.integer(N))
```

Plot
```{r}
recode_vec = c("Infertile male",
               "No males",
               "No females", 
               "Not productive",
               "All-cause death", 
               "Reproducing")
names(recode_vec) = c("infertile male", 
                      "no males", 
                      "no females", 
                      "not productive", 
                      "all cause death", 
                      "N_SURVIVING")

test = inbreed_df_tidy %>% 
  # geom_area() seems to need `x` to be numeric
  dplyr::mutate(GENERATION = as.integer(gsub("F", "", GENERATION))) %>%
  # recode STATUS values and order
  dplyr::mutate(STATUS = dplyr::recode(STATUS, !!!recode_vec),
                STATUS = factor(STATUS, levels = recode_vec))# %>% 
out = ggplot(test) +
    geom_area(aes(x=GENERATION,
                  y=N,
                  fill = STATUS)) +
#    geom_text(aes(x=GENERATION, y=N, label = STATUS)) +
    scale_x_continuous(breaks = unique(test$GENERATION),
                       labels = paste("F", unique(test$GENERATION), sep = "")) +
    theme_bw() +
    xlab("Generation") +
    ylab("Number of MIKK panel lines") +
    labs(fill = "Status") +
    theme(panel.grid = element_blank())
#    theme(legend.position=c(0.2, 0.3),
#          legend.box.background = element_blank())

out
```
```{r, eval =F}
ggsave(here("plots", "20210215_inbreeding_status.svg"),
       device = "svg",
       units = "cm",
       dpi = 400,
       height = 12,
       width = 19.5)
```


```{r}
ggplotly(out)
```

# Compile final figure

## Create new plots with new palettes

### Inbreeding

```{r}
recode_vec = c("Infertile male",
               "No males",
               "No females", 
               "Not productive",
               "All-cause death", 
               "Reproducing")
names(recode_vec) = c("infertile male", 
                      "no males", 
                      "no females", 
                      "not productive", 
                      "all cause death", 
                      "N_SURVIVING")

test = inbreed_df_tidy %>% 
  # geom_area() seems to need `x` to be numeric
  dplyr::mutate(GENERATION = as.integer(gsub("F", "", GENERATION))) %>%
  # recode STATUS values and order
  dplyr::mutate(STATUS = dplyr::recode(STATUS, !!!recode_vec),
                STATUS = factor(STATUS, levels = recode_vec))# %>% 
out2 = ggplot(test) +
    geom_area(aes(x=GENERATION,
                  y=N,
                  fill = STATUS)) +
#    geom_text(aes(x=GENERATION, y=N, label = STATUS)) +
    scale_fill_viridis_d(option = "magma") +
    scale_x_continuous(breaks = unique(test$GENERATION),
                       labels = paste("F", unique(test$GENERATION), sep = "")) +
    theme_cowplot(font_size = 10) +
    xlab("Generation") +
    ylab("Number of MIKK panel lines") +
    labs(fill = "Status") +
    theme(panel.grid = element_blank())
#    theme(legend.position=c(0.2, 0.3),
#          legend.box.background = element_blank())

out2
```

### Fecundity horizontal

```{r}
# Process data
final_hor_df = semi_out %>% 
  dplyr::mutate(KEY = gsub("; ", ";\n", KEY),
                KEY = factor(KEY, levels = recode_vec_3)) %>% 
  dplyr::filter(DATE == "Jul 2020") %>% 
  dplyr::mutate(STRAIN = factor(STRAIN, levels=unique(STRAIN[order(KEY,decreasing = T)])))
  # order by category

# Get levels for correlation plot
pal_corr = viridis_pal(direction = -1)(length(levels(final_hor_df$STRAIN)))
names(pal_corr) = levels(final_hor_df$STRAIN)

final_hor_plot = final_hor_df %>%   
  ggplot() +
    geom_col(aes(STRAIN, KEY, fill = STRAIN)) +
    theme_cowplot() +
    scale_fill_viridis_d(direction = -1) +
    guides(fill = F) +
    theme(axis.text.x = element_text(size = 3.5,
                                     angle = 45),
          axis.text.y = element_text(size = 5),
          axis.title.x = element_text(size = 9)) +
    xlab("F16 MIKK panel lines (July 2020)") +
    ylab(NULL)

final_hor_plot
```
### Correlation plot

```{r}
recode_vec_4 = c("Not producing",
                 "Do not produce\nevery day;\n<3 eggs\nwhen they do",
                 "Do not produce\nevery day;\n<5 eggs\nwhen they do",
                 "0-3 eggs\nper day",
                 "0-5 eggs\nper day",
                 "5-10 eggs\nper day")
names(recode_vec_4) = recode_vec_2
  
  
semi_corr = semi_out %>% 
  dplyr::mutate(KEY = dplyr::recode_factor(KEY, !!!recode_vec_4)) %>% 
  tidyr::pivot_wider(id_cols = STRAIN,
                     names_from = DATE,
                     values_from = KEY) %>% 
  ggplot() +
    geom_jitter(aes(`Feb 2019`, `Jul 2020`, colour = STRAIN), alpha = 0.8) +
    coord_fixed() +
    theme_cowplot() +
    theme(axis.text = element_text(size = 5),
          axis.title = element_text(size = 9)) +
    guides(colour = F) +
    scale_color_manual(values = pal_corr) +
    xlab("F15: Feb 2019") +
    ylab("F16: Jul 2020")

semi_corr
```


```{r}
ggdraw() +
  draw_plot(out2, x = 0.05, y = .5, width = .45, height = .5) +
  draw_plot(semi_corr, x = .4, y = .35, width = .65, height = .65) +
  draw_plot(final_hor_plot, x = 0, y = 0, width = 1, height = 0.5) +
  draw_plot_label(label = c("A", "B", "C"), size = 15,
                  x = c(0, 0.5, 0), y = c(1, 1, 0.5))
```

```{r, eval = F}
# SVG
ggsave(here("plots", "fecundity", "20210217_final_figure.svg"),
       device = "svg",
       units = "cm",
       height = 15,
       width = 30)

# PNG
ggsave(here("plots", "fecundity", "20210217_final_figure.png"),
       device = "png",
       dpi = 400,
       units = "cm",
       height = 15,
       width = 30)
```

```{julia}
2 + 2
```

