---
title: "Structural variants"
author: "Ian Brettell"
date: '`r format(Sys.Date())`'
output:
  html_document:
    toc: true
    toc_float: true
    dev: 'svg'
    number_sections: true
    keep_md: true
    pandoc_args: --lua-filter=color-text.lua
    highlight: pygments  
---

Email from Tom, 18 January:

[1.]{color="purple"} “Polish” SV calls using Illumina data:
<https://github.com/smehringer/SViper>

Please explore the above - get install and tested.

Nanopore VCF files:

* [WITH SEQUENCES]{color="red"}:
`ll /hps/research1/birney/users/adrien/indigene/analyses/indigene_nanopore_DNA/brain_run2/DNA_analysis/results/SV/sniffles_all_OLD/merged.vcf`
* [WITHOUT SEQUENCES]{color="red"}:`/hps/research1/birney/users/adrien/indigene/analyses/indigene_nanopore_DNA/brain_run2/DNA_analysis/results/SV/sniffles_all/merged.vcf`

`SViper` says:
>The vcf file must be a structural variant format (tags instead of sequences, e.g. <DEL>). ALso the INFO field must include the END tag, giving the end position of the variant, as well as the SVLEN tag in case of insertions.

Then we need to map Illumina ID to the Nanopore VCF and figure out how to run the polishing.

Once done we need to summarise and describe the SV’s a little with some annotations etc
- i will make a start with doing this part and we catchup once you have looked at the polishing

[2.]{color="purple"} catchup with Jack on repeats (both MIKK panel and Nanopore assemblies).
- create one plot for each paper

3. Fecundity - i will send you the data and a description of what exactly it is later today.
- we will need a general description of the data, descriptive stats and a heritability estimate - plus a single plot or table (for MIKK panel paper)

4. Add to txt within the two documents - we can catch-up on this later in the week.

# Polish SV calls with Illumina data

## Setup

Working directory on EBI cluster: `/hps/research1/birney/users/ian/mikk_paper/mikk_genome/sv_analysis`
GitHub repo: <https://github.com/brettellebi/mikk_genome>

### `conda` envs

#### `sv_env`

```{bash, eval = F}
# Make conda env
mamba create -n sv_env
conda activate sv_env
# Install packages
mamba install -c bioconda bcftools sniffles
mamba install -c conda-forge r-base r-tidyverse
# Export to file
conda env export > envs/sv_env/sv_env.yaml
```

#### `baseR`

```{bash, eval = F}
# Activate
conda activate baseR
# Export
conda env export > envs/baseR/baseR.yaml
```

### `renv`

```{r, eval = F}
# Initiate
renv::init()
# Snapshot
renv::snapshot()
# Restore
renv::restore()
```

#### Source

```{r}
library(here)
source(here::here("code", "scripts", "sv_analysis", "source.R"))
```

### Nanopore VCF

`/hps/research1/birney/users/adrien/analyses/medaka_DNA_promethion/brain_run2/DNA_analysis/results/SV/sniffles_all/merged.vcf`

Created with:
```{bash}
# First used:
sniffles \
  --min_support 3 \
  --max_num_splits 7 \
  --max_distance 1000 \
  --min_length 50 \
  --minmapping_qual 20 \
  --min_seq_size 1000 \
  --allelefreq 0.1 \
  -t {threads} \
  -m {input_bam} \
  -v {output_vcf}
# Then I filtered and merged all the variants from the different samples together with survivor and recalled variants a second time in forced mode using the merged set with sniffles again using the same options.
```

#### Copy to working directory

```{bash, eval = F}
# With sequences
nano_raw=/hps/research1/birney/users/adrien/indigene/analyses/indigene_nanopore_DNA/brain_run2/DNA_analysis/results/SV/sniffles_all_OLD/merged.vcf
out_path=../sv_analysis/vcfs/ont_raw_with_seq.vcf
## Copy
cp $nano_raw $out_path

# Without sequences
nano_raw=/hps/research1/birney/users/adrien/indigene/analyses/indigene_nanopore_DNA/brain_run2/DNA_analysis/results/SV/sniffles_all/merged.vcf
out_path=../sv_analysis/vcfs/ont_raw.vcf
## Copy
cp $nano_raw $out_path
```

#### Rename samples

##### With sequences

```{bash, eval = F}
conda activate sv_env

in_vcf=../sv_analysis/vcfs/ont_raw_with_seq.vcf
sample_file=data/sv_analysis/20210205_ont_raw_samples_file.txt
out_vcf=../sv_analysis/vcfs/ont_raw_with_seq_rehead.vcf

# Make samples key file
bcftools query -l $in_vcf \
  > tmp1
cut -f4 -d'/' tmp1 | cut -f1 -d'_' \
  > tmp2
paste -d' ' tmp1 tmp2 > $sample_file
rm tmp1 tmp2

# Rename VCF
bcftools reheader \
  --samples $sample_file \
  --output $out_vcf \
  $in_vcf
```

##### Without sequences

```{bash, eval = F}
conda activate sv_env

in_vcf=../sv_analysis/vcfs/ont_raw.vcf
sample_file=data/sv_analysis/20210212_ont_raw_samples_file.txt
out_vcf=../sv_analysis/vcfs/ont_raw_rehead.vcf

# Make samples key file
bcftools query -l $in_vcf \
  > tmp1
cut -f4 -d'/' tmp1 | cut -f1 -d'_' \
  > tmp2
paste -d' ' tmp1 tmp2 > $sample_file
rm tmp1 tmp2

# Rename VCF
bcftools reheader \
  --samples $sample_file \
  --output $out_vcf \
  $in_vcf
```

#### Get stats

##### With sequences

```{bash, eval = F}
conda activate sv_mikk

in_vcf=../sv_analysis/vcfs/ont_raw_with_seq.vcf
stats_out=../sv_analysis/vcfs/ont_raw_with_seq.stats

# Get stats
bcftools stats \
  $in_vcf \
    > $stats_out
```

##### Without sequences

```{bash, eval = F}
conda activate sv_mikk

in_vcf=../sv_analysis/vcfs/ont_raw.vcf
stats_out=../sv_analysis/vcfs/ont_raw.stats

# Get stats
bcftools stats \
  $in_vcf \
    > $stats_out
```

#### Split per sample

##### With sequences

```{bash}
conda activate sv_mikk

in_vcf=../sv_analysis/vcfs/ont_raw_with_seq_rehead.vcf
out_dir=../sv_analysis/vcfs/ont_raw_with_seq_rehead_per_sample

mkdir -p $out_dir

# Split by sample
bcftools +split \
  $in_vcf \
  --output $out_dir
```

##### Without sequences

```{bash}
conda activate sv_mikk

in_vcf=../sv_analysis/vcfs/ont_raw_rehead.vcf
out_dir=../sv_analysis/vcfs/ont_raw_rehead_per_sample

mkdir -p $out_dir

# Split by sample
bcftools +split \
  $in_vcf \
  --output $out_dir
```

### Illumina VCF

#### Copy to working directory

```{bash, eval = F}
conda activate sv_env

in_vcf=/nfs/research1/birney/projects/medaka/inbred_panel/medaka-alignments-release-94/vcf/medaka_inbred_panel_ensembl_new_reference_release_94.vcf
out_vcf=../sv_analysis/vcfs/ill_raw.vcf.gz

# Compress and copy
bsub \
  -M 30000 \
  -o ../log/20210208_comp_ill.out \
  -e ../log/20210208_comp_ill.err \
bsub -Is bash \
  """
  conda activate sv_env ;
  bcftools view \
    --output-type z \
    --output $out_vcf \
    $in_vcf
  """  
```

#### Rename and filter for ONT samples

##### Pull out IDs for relevant samples

```{r, eval = F}
library(here)
library(tidyverse)
ont_samples = here::here("data", "sv_analysis", "20210205_ont_raw_samples_file.txt")
ill_samples = here::here("data","20200206_cram_id_to_line_id.txt")
out_file = here::here("data", "sv_analysis", "20210205_ill_key_ont_samples.txt")
out_samples = here::here("data", "sv_analysis", "20210205_ont_samples_only.txt")

# Read in tables

ont_key = read.table(ont_samples)
ill_key = read.table(ill_samples, comment.char = "\"", header = T) %>% 
  dplyr::mutate(line = gsub("_", "-", line))

# Find matches
out = ill_key[ill_key$line %in% ont_key$V2, ]

# Write to files
## Key file
readr::write_delim(out, out_file, delim = " ", col_names = F)
## Just samples
readr::write_lines(out$cram_file, out_samples)
```

##### Rename and filter

```{bash, eval = F}
in_vcf=../sv_analysis/vcfs/ill_raw.vcf.gz
samples_file=data/sv_analysis/20210205_ont_samples_only.txt
samples_key=data/sv_analysis/20210205_ill_key_ont_samples.txt
out_vcf=../sv_analysis/vcfs/ill_raw_rehead.vcf.gz
out_dir=../sv_analysis/vcfs/ill_raw_rehead_per_sample

mkdir -p $out_dir

# Filter for target samples and rehead
bcftools view \
  --samples-file $samples_file \
  --output-type u \
  $in_vcf |\
    bcftools reheader \
      --samples $samples_key \
      --output $out_vcf

# Split by sample
bcftools +split \
  $out_vcf \
  --output $out_dir
```

### BAMs

#### Illumina `.bam` files

Copy to local.

```{bash}
sample_key=data/sv_analysis/20210205_ill_key_ont_samples.txt
ill_bam_dir=/nfs/research1/birney/projects/medaka/inbred_panel/medaka-alignments-release-94/bam
out_dir=../sv_analysis/bams

mkdir -p $out_dir 

# Copy over
for sample in $(cut -f1 -d' ' $sample_key ) ; do
  cp $ill_bam_dir/$sample.bai $out_dir ;
done  

# SViper needs bams in .bam.bai format. Original {sample}.bai files need to be copied to {sample}.bam.bai
for file in $( find $out_dir/*.bai ) ; do
  new_filename=$( echo $file | sed 's/.bai/.bam.bai/g' ) ;
  mv $file $new_filename ;
done
```

#### Nanopore `.bam` files

Sit here: `/hps/research1/birney/users/adrien/indigene/analyses/indigene_nanopore_DNA/brain_run2/DNA_analysis/results/SV/ngmlr_alignments/`

## Polish Nanopore reads with `SViper`

### Create Singularity container
```{bash}
module load singularity

# Build
singularity build \
  --remote ../sing_conts/sviper.sif \
  envs/sviper/20210204_sviper.def
  
# Open interactive shell
bsub -Is "singularity shell ../sing_conts/sviper.sif"
# Works! 
```

### Test

```{bash}
# Load singularity
module load singularity
# Pull image built with `envs/sviper/20210204_sviper.def`
bsub -M 30000 -n 4 -Is "singularity shell ../sing_conts/sviper.sif"

#######################
# Variables
#######################
## Container
container=../sing_conts/sviper.sif
## Sample
sample=11-1
sample_key=data/sv_analysis/20210205_ill_key_ont_samples.txt
## VCF to polish
ont_vcf=../sv_analysis/vcfs/ont_raw_rehead_per_sample/$sample.vcf
## Illumina BAM
ill_cram_id=$(grep $sample $sample_key | cut -f1 -d' ')
ill_bam_dir=../sv_analysis/bams
ill_bam=$ill_bam_dir/$ill_cram_id.bam
## Nanopore BAM
ont_bam_dir=/hps/research1/birney/users/adrien/indigene/analyses/indigene_nanopore_DNA/brain_run2/DNA_analysis/results/SV/ngmlr_alignments
ont_bam=$(find $ont_bam_dir/$sample*.bam)
## Reference
ref=../refs/Oryzias_latipes.ASM223467v1.dna.toplevel.fa
## Output directory
out_dir=../sv_analysis/vcfs/sviper
mkdir -p $out_dir

#######################
# TEST call sviper 
#######################
sviper \
  --candidate-vcf $ont_vcf \
  --short-read-bam $ill_bam \
  --long-read-bam $ont_bam \
  --reference $ref \
  --output-prefix $out_dir/$sample

```

### True

```{bash}
# Load singularity
module load singularity
# Pull image built with `envs/sviper/20210204_sviper.def`
bsub -M 30000 -n 4 -Is "singularity shell ../sing_conts/sviper.sif"

#######################
# Variables
#######################
## Sample
sample=11-1
sample_key=data/sv_analysis/20210205_ill_key_ont_samples.txt
## VCF to polish
ont_vcf=../sv_analysis/vcfs/ont_raw_rehead_per_sample/$sample.vcf
## Illumina BAM
ill_cram_id=$(grep $sample $sample_key | cut -f1 -d' ')
ill_bam_dir=../sv_analysis/bams
ill_bam=$ill_bam_dir/$ill_cram_id.bam
## Nanopore BAM
ont_bam_dir=/hps/research1/birney/users/adrien/indigene/analyses/indigene_nanopore_DNA/brain_run2/DNA_analysis/results/SV/ngmlr_alignments
ont_bam=$(find $ont_bam_dir/$sample*.bam)
## Reference
ref=../refs/Oryzias_latipes.ASM223467v1.dna.toplevel.fa
## Container
container=../sing_conts/sviper.sif
## Output directory
out_dir=../sv_analysis/vcfs/sviper
mkdir -p $out_dir

#######################
# TEST call sviper 
#######################
sviper \
  --candidate-vcf $ont_vcf \
  --short-read-bam $ill_bam \
  --long-read-bam $ont_bam \
  --reference $ref \
  --output-prefix $out_dir/$sample
  
#######################
# TRUE call sviper 
####################### 
module load singularity

# Global variables
## Sample key
sample_key=data/sv_analysis/20210205_ill_key_ont_samples.txt
## BAM dirs
ill_bam_dir=../sv_analysis/bams
ont_bam_dir=/hps/research1/birney/users/adrien/indigene/analyses/indigene_nanopore_DNA/brain_run2/DNA_analysis/results/SV/ngmlr_alignments
## Reference
ref=../refs/Oryzias_latipes.ASM223467v1.dna.toplevel.fa
## Container
container=../sing_conts/sviper.sif
## Output directory
out_dir=../sv_analysis/vcfs/sviper
mkdir -p $out_dir

for sample in $(cut -f2 -d' ' $sample_key | tail -n+2 ) ; do
  # Set variables
  
  ## VCF to polish
  ont_vcf=../sv_analysis/vcfs/ont_raw_rehead_per_sample/$sample.vcf
  ## Illumina BAM
  ill_cram_id=$(grep " $sample" $sample_key | cut -f1 -d' ')
  ill_bam=$ill_bam_dir/$ill_cram_id.bam
  ## Nanopore BAM
  ont_bam=$(find $ont_bam_dir/$sample*.bam)

  # Run SViper
  bsub \
    -M 30000 \
    -n 4 \
    -o ../log/20210212_sviper_$sample.out \
    -e ../log/20210212_sviper_$sample.err \
    """
    singularity exec $container \
      sviper \
        --candidate-vcf $ont_vcf \
        --short-read-bam $ill_bam \
        --long-read-bam $ont_bam \
        --reference $ref \
        --output-prefix $out_dir/$sample
    """
done
```

