---
title: "Nucleotide diversity"
date: '`r format(Sys.Date())`'
#output: html_notebook
#editor_options: 
#  chunk_output_type: inline
output:
  html_document:
    toc: true
    toc_float: true
    dev: 'svg'
    number_sections: true
    keep_md: false
    pandoc_args: --lua-filter=color-text.lua
    highlight: pygments 
---

# Load libraries

```{r, message = F, warning = F}
library(here)
library(tidyverse)
library(karyoploteR)
library(plotly)
library(circlize)
library(GGally)
library(viridis)
```


# Snakemake pipeline to process VCF and extract data

<https://github.com/brettellebi/mikk_genome/tree/master/code/snakemake/nucleotide_diversity>

**Steps**

1. Filter MIKK panel VCF for non-sibling lines (N = 63).
1. Use `VCFtools`'s `--window-pi` to calculate nucleotide divergence in different-sized windows.
1. Use `bcftools` to extract `MQ` from `INFO` field.

# Process data

Data location: `/nfs/research/birney/users/ian/mikk_genome/nucleotide_divergence`

## Read in data

### MIKK

```{r}
# Set location of data
in_dir = "/nfs/research/birney/users/ian/mikk_genome/nucleotide_divergence/mikk/all"

in_files = list.files(in_dir, full.names = T)
dat_list = lapply(in_files, function(IN_FILE) {
  # Read in 
  readr::read_delim(IN_FILE,
                    delim = "\t",
                    col_types = c("ciiid")) %>% 
    # Remove MT
    dplyr::filter(CHROM != "MT") %>% 
    # Make CHR an integer
    dplyr::mutate(CHROM = as.integer(CHROM)) %>% 
    # Create middle of bin
    dplyr::mutate(BIN_MID = ((BIN_END - BIN_START - 1) / 2) + BIN_START )
})
names(dat_list) = basename(in_files) %>%
  str_remove(".windowed.pi")
```

### Wild Kiyosu

```{r}
# Set location of data
in_dir = "/nfs/research/birney/users/ian/mikk_genome/nucleotide_divergence/wild_kiyosu"

in_files = list.files(in_dir, full.names = T)
wk_list = lapply(in_files, function(IN_FILE) {
  # Read in 
  readr::read_delim(IN_FILE,
                    delim = "\t",
                    col_types = c("ciiid")) %>% 
    # Remove MT
    dplyr::filter(CHROM != "MT") %>% 
    # Make CHR an integer
    dplyr::mutate(CHROM = as.integer(CHROM)) %>% 
    # Create middle of bin
    dplyr::mutate(BIN_MID = ((BIN_END - BIN_START - 1) / 2) + BIN_START )
})
names(wk_list) = basename(in_files) %>%
  str_remove(".windowed.pi")
```

# Karyoplot

## Set up scaffold

```{r}
med_chr_lens = read.table(here::here("data",
                                     "Oryzias_latipes.ASM223467v1.dna.toplevel.fa_chr_counts.txt"),
                          col.names = c("chr", "end"))
# Add start
med_chr_lens$start = 1
# Reorder
med_chr_lens = med_chr_lens %>% 
  dplyr::select(chr, start, end) %>% 
  # remove MT
  dplyr::filter(chr != "MT") %>% 
  # convert to numeric
  dplyr::mutate(chr = as.integer(chr)) %>% 
  # order
  dplyr::arrange(chr)

# Create custom genome 
med_genome = regioneR::toGRanges(med_chr_lens)
```

## Plot

```{r, fig.dim=c(10, 3)}
# Test with ggplot
dat_list$`1000000` %>% 
  ggplot() +
    geom_line(aes(BIN_MID, PI)) +
    facet_grid(cols = vars(CHROM))

dat_list$`500000` %>% 
  ggplot() +
    geom_line(aes(BIN_MID, PI)) +
    facet_grid(cols = vars(CHROM))

dat_list$`100000` %>% 
  ggplot() +
    geom_line(aes(BIN_MID, PI)) +
    facet_grid(cols = vars(CHROM))
```

```{r}
## Get max Y
round.choose <- function(x, roundTo, dir = 1) {
  if(dir == 1) {  ##ROUND UP
    x + (roundTo - x %% roundTo)
  } else {
    if(dir == 0) {  ##ROUND DOWN
      x - (x %% roundTo)
    }
  }
}
y_max = round.choose(max(dat_list$`1000000`$PI), 0.01, 1)
```


```{r, eval = F}
# set file name
file_name = paste("20210930_", "1Mb", ".png", sep = "")
file_out = here::here("docs/plots/nucleotide_diversity", file_name)

png(file=file_out,
    width=13000,
    height=1000,
    units = "px",
    res = 300)

# Plot ideogram
kp = karyoploteR::plotKaryotype(med_genome, plot.type = 5)

# Add base numbers 
karyoploteR::kpAddBaseNumbers(kp, tick.dist = 5000000, cex = 0.7)

# Set y-axis limits
karyoploteR::kpAxis(kp, ymin=0, ymax=y_max )

# Add lines
lwd = 1
karyoploteR::kpLines(kp,
                     chr = dat_list$`1000000`$CHROM,
                     x = dat_list$`1000000`$BIN_MID,
                     y = dat_list$`1000000`$PI,
                     ymax = y_max,
                     r0=0, r1 = 1,
                     lwd = lwd)



dev.off()  
```

```{r}
file_out = here::here("docs/plots/nucleotide_diversity/20210930_1Mb.png")
knitr::include_graphics(file_out)
```

# Circos

## Get repeats data

```{r, message = F}
repeats_file = "/nfs/research/birney/users/ian/mikk_genome/repeats/medaka_hdrr_repeats.fixed.gff"

hdrr_reps = readr::read_delim(repeats_file,
                              delim = "\t",
                              col_names = F,
                              skip = 3,
                              comment = "",
                              quote = "") %>%
  # Remove empty V8 column
  dplyr::select(-X8) %>% 
  # Get class of repeat from third column
  dplyr::mutate(class = stringr::str_split(X3, pattern = "#", simplify = T)[, 1]) %>% 
  # Rename columns
  dplyr::rename(chr = X1, tool = X2, class_full = X3, start = X4, end = X5, percent = X6, strand = X7, info = X9)

# Find types of class other than "(GATCCA)n" types
class_types = unique(hdrr_reps$class[grep(")n", hdrr_reps$class, invert = T)])

hdrr_reps = hdrr_reps %>%
  # NA for blanks
  dplyr::mutate(class = dplyr::na_if(class, "")) %>%
  # "misc" for others in "(GATCCA)n" type classes
  dplyr::mutate(class = dplyr::if_else(!class %in% class_types, "Miscellaneous", class)) %>%
  # rename "Simple_repeat"
  dplyr::mutate(class = dplyr::recode(class, "Simple_repeat" = "Simple repeat")) %>% 
  # filter out NA in `chr` (formerly MT)
  dplyr::filter(!is.na(chr))


```

## Calculate proportion of bin covered by repeats

### Bin intevals

```{r}
# Create GRanges object with bins
bin_length = "1000000"

bin_intervals = dat_list[[bin_length]] %>% 
  dplyr::select(CHROM, BIN_START, BIN_END) %>% 
  # split into list by chromosome
  split(., f = .$CHROM)

# Replace END pos of final bin for each chromsome with actual end pos
bin_intervals = lapply(1:length(bin_intervals), function(CHR){
  # Get end position for target chromosome
  end_pos = med_chr_lens %>% 
    dplyr::filter(chr == CHR) %>% 
    dplyr::pull(end)
  # Replace final bin end position
  out = bin_intervals[[CHR]]
  out$BIN_END[nrow(out)] = end_pos
  
  return(out)
}) %>% 
  dplyr::bind_rows()

# Convert to data frame
bin_intervals = as.data.frame(bin_intervals)

# Convert to GRanges
bin_ranges = regioneR::toGRanges(bin_intervals)
```

### Repeats

```{r}
# Convert `hdrr_reps` to GRanges
hdrr_ranges = GenomicRanges::makeGRangesFromDataFrame(hdrr_reps,
                                                      keep.extra.columns = T,
                                                      seqnames.field = "chr",
                                                      start.field = "start",
                                                      end.field = "end")

# Get non-overlapping regions
hdrr_covered = GenomicRanges::disjoin(hdrr_ranges)
```

### Find quantity of repeats overlapping bins

```{r}
overlaps = GenomicRanges::findOverlaps(bin_ranges, hdrr_covered)

# Split into list by bin
overlaps_list = lapply(unique(overlaps@from), function(BIN){
  out = list()
  # Get indexes of all repeat ranges overlapping the target BIN
  out[["hits"]] = overlaps[overlaps@from == BIN]@to
  # Extract those ranges from `hdrr_ranges`
  out[["range_hits"]] = hdrr_covered[out[["hits"]]]
    # Get number bases covered by each range
  out[["widths"]] = out[["range_hits"]] %>% 
    GenomicRanges::width(.) 
  # Get summed widths
  out[["summed"]] = out[["widths"]] %>% 
    # Get total bases covered in bin
    sum(.)
  
  return(out)
})
  
overlaps_vec = purrr::map(overlaps_list, function(BIN){
  BIN[["summed"]]
}) %>% 
  unlist(.)

# Add as column to `bin_intervals`
bin_intervals$REPEAT_COV = overlaps_vec

# Caclulate proportion
bin_intervals = bin_intervals %>%
  dplyr::mutate(REPEAT_PROP = REPEAT_COV / (BIN_END - BIN_START + 1))
```

## Read in mapping quality scores

```{r}
in_file = "/nfs/research/birney/users/ian/mikk_genome/mapping_quality/mapping_quality.csv"

mq_df = readr::read_csv(in_file,
                        col_names = c("CHROM", "POS", "MQ"),
                        col_types = c("cid")) %>% 
  # remove 'MT'
  dplyr::filter(!CHROM == "MT") %>% 
  # make CHROM integer
  dplyr::mutate(CHROM = as.integer(CHROM))

# Bin and get means
mq_list = mq_df %>% 
  split(., f = .$CHROM)

# Bin into 1 Mb intervals
binned_mq_list = purrr::map(mq_list, function(CHR){
  # Set intervals
  intervals = seq(1, max(CHR$POS), by = 1000000)
  # add final length
  if (max(intervals) != max(CHR$POS)) {
    intervals = c(intervals, max(CHR$POS))
  }
  # bin
  CHR = CHR %>% 
    dplyr::mutate(BIN = cut(POS,
                            breaks = intervals,
                            labels = F,
                            include.lowest = T))
})

# calculate mean MQ within each bin
mean_mq = purrr::map(binned_mq_list, function(CHR){
  CHR %>% 
    # replace Inf with NA
    dplyr::mutate(MQ = dplyr::na_if(MQ, Inf)) %>% 
    # Group by BIN
    dplyr::group_by(BIN) %>% 
    # Calculate mean MQ within each bin
    dplyr::summarise(MEAN_MQ = mean(MQ, na.rm = T))
}) %>% 
  # bind into single data frame
  dplyr::bind_rows(.id = "CHR")


```

## Bind data frames

```{r}
# Bind data frames
pi_repeat_df = cbind(dat_list$`1000000`,
                     PI_WK = wk_list[["1000000"]]$PI,
                     REPEAT_PROP = bin_intervals$REPEAT_PROP,
                     MEAN_MQ = mean_mq$MEAN_MQ)
```

## Pairs plot

```{r}
pairs_plot = GGally::ggpairs(pi_repeat_df, columns = c(5, 7, 8, 9),
                             lower = list(continuous = wrap("points", alpha = 0.2, size = 0.5),
                                          combo = wrap("dot_no_facet", alpha = 0.4)))

plotly::ggplotly(pairs_plot)
```


## Circos

### 1 Mb

#### Rearrange data frames

```{r}
df_circos_1Mb = pi_repeat_df
# take `BIN_END` from `bin_intervals` with chromosome end number for final bin
df_circos_1Mb$BIN_END = bin_intervals$BIN_END
```

#### Plot

```{r}
out_plot = here::here("docs/plots/nucleotide_diversity", "20211001_circos_1Mb.png")
```

```{r}
png(out_plot,
    width = 20,
    height = 20,
    units = "cm",
    res = 500)

# Set parameters
## Decrease cell padding from default c(0.02, 1.00, 0.02, 1.00)
circos.par(cell.padding = c(0, 0, 0, 0),
           track.margin = c(0, 0),
           gap.degree = c(rep(1, nrow(med_chr_lens) - 1), 6))
# Initialize plot
circos.initializeWithIdeogram(med_chr_lens,
                              plotType = c("axis", "labels"),
                              major.by = 1e7,
                              axis.labels.cex = 0.25*par("cex"))

# Add MIKK Pi line
circos.genomicTrack(df_circos_1Mb,
    panel.fun = function(region, value, ...){
  circos.genomicLines(region,
                      value[[2]],
                      col = "#49A379",
                      area = T,
                      border = karyoploteR::darker("#49A379"))
    },
    track.height = 0.1,
    bg.border = NA,
    ylim = c(0, 0.021))

circos.yaxis(side = "right",
           at = c(.01, .02),
           labels.cex = 0.25*par("cex"),
           tick.length = 2
           )

circos.text(0, 0.01,
        labels = expression(paste("MIKK (", pi, ")", sep = "")),
        sector.index = "1",
        facing = "clockwise",
        adj = c(.5, 0),
        cex = 0.4*par("cex"))  

# Add wild Kiyosu line
circos.genomicTrack(df_circos_1Mb,
    panel.fun = function(region, value, ...){
  circos.genomicLines(region,
                      value[[4]],
                      col = "#7A306C",
                      area = T,
                      border = karyoploteR::darker("#8A4F7D"))
    },
    track.height = 0.1,
    bg.border = NA,
    ylim = c(0, 0.021))

circos.yaxis(side = "right",
           at = c(.01, .02),
           labels.cex = 0.25*par("cex"),
           tick.length = 2
           )

circos.text(0, 0.01,
        labels = expression(paste("Wild\nKiyosu (", pi, ")", sep = "")),
        sector.index = "1",
        facing = "clockwise",
        adj = c(.5, 0),
        cex = 0.4*par("cex"))  

# Add repeat content
circos.genomicTrack(df_circos_1Mb,
  panel.fun = function(region, value, ...) {
    circos.genomicLines(region,
                        value[[5]],
                        type = "h",
                        col = "#0C1B33",
                        cex = 0.05,
                        baseline = 0)
  },
  track.height = 0.1,
  ylim = c(0,0.5),
  bg.border = NA)

circos.yaxis(side = "right",
           at = c(0, .5, 1),
           labels.cex = 0.25*par("cex"),
           tick.length = 5
           )

circos.text(0, 0.25,
        labels = "Repeat\ncontent",
        sector.index = "1",
        facing = "clockwise",
        adj = c(.5, 0),
        cex = 0.4*par("cex")) 

# Add mean mapping quality
circos.genomicTrack(df_circos_1Mb,
  panel.fun = function(region, value, ...) {
    circos.genomicLines(region,
                        value[[6]],
                        type = "h",
                        col = "#FF6978",
                        cex = 0.05,
                        baseline = 20)
  },
  track.height = 0.1,
  ylim = c(20,65),
  bg.border = NA)

circos.yaxis(side = "right",
           at = c(20, 50),
           labels.cex = 0.25*par("cex"),
           tick.length = 5
           )

circos.text(0, 40,
        labels = "Mean MQ",
        sector.index = "1",
        facing = "clockwise",
        adj = c(.5, 0),
        cex = 0.4*par("cex")) 

# Add baseline
#circos.xaxis(h = "bottom",
#             labels = F,
#             major.tick = F)

# Print label in center
text(0, 0, "Nucleotide diversity in\nMIKK and wild Kiyosu medaka,\nrepeat content,\nand mapping quality\nin 1 Mb windows")

circos.clear()

dev.off()
```

```{r}
knitr::include_graphics(out_plot)
```

### 500 Kb

#### Create data frame

##### Repeats

```{r}
# Create GRanges object with bins
bin_length = "500000"

bin_intervals_500 = dat_list[[bin_length]] %>% 
  dplyr::select(CHROM, BIN_START, BIN_END) %>% 
  # split into list by chromosome
  split(., f = .$CHROM)

# Replace END pos of final bin for each chromsome with actual end pos
bin_intervals_500 = lapply(1:length(bin_intervals_500), function(CHR){
  # Get end position for target chromosome
  end_pos = med_chr_lens %>% 
    dplyr::filter(chr == CHR) %>% 
    dplyr::pull(end)
  # Replace final bin end position
  out = bin_intervals_500[[CHR]]
  out$BIN_END[nrow(out)] = end_pos
  
  return(out)
}) %>% 
  dplyr::bind_rows()

# Convert to data frame
bin_intervals_500 = as.data.frame(bin_intervals_500)

# Convert to GRanges
bin_ranges = regioneR::toGRanges(bin_intervals_500)
  
overlaps = GenomicRanges::findOverlaps(bin_ranges, hdrr_covered)

# Split into list by bin
overlaps_list = lapply(unique(overlaps@from), function(BIN){
  out = list()
  # Get indexes of all repeat ranges overlapping the target BIN
  out[["hits"]] = overlaps[overlaps@from == BIN]@to
  # Extract those ranges from `hdrr_ranges`
  out[["range_hits"]] = hdrr_covered[out[["hits"]]]
    # Get number bases covered by each range
  out[["widths"]] = out[["range_hits"]] %>% 
    GenomicRanges::width(.) 
  # Get summed widths
  out[["summed"]] = out[["widths"]] %>% 
    # Get total bases covered in bin
    sum(.)
  
  return(out)
})
  
overlaps_vec = purrr::map(overlaps_list, function(BIN){
  BIN[["summed"]]
}) %>% 
  unlist(.)

# Add as column to `bin_intervals_500`
bin_intervals_500$REPEAT_COV = overlaps_vec

# Caclulate proportion
bin_intervals_500 = bin_intervals_500 %>%
  dplyr::mutate(REPEAT_PROP = REPEAT_COV / (BIN_END - BIN_START + 1))
```

##### MQ

```{r}
# Bin into 1 Mb intervals
binned_mq_list_500 = purrr::map(mq_list, function(CHR){
  # Set intervals
  intervals = seq(1, max(CHR$POS), by = 500000)
  # add final length
  if (max(intervals) != max(CHR$POS)) {
    intervals = c(intervals, max(CHR$POS))
  }
  # bin
  CHR = CHR %>% 
    dplyr::mutate(BIN = cut(POS,
                            breaks = intervals,
                            labels = F,
                            include.lowest = T))
})

# calculate mean MQ within each bin
mean_mq_500 = purrr::map(binned_mq_list_500, function(CHR){
  CHR %>% 
    # replace Inf with NA
    dplyr::mutate(MQ = dplyr::na_if(MQ, Inf)) %>% 
    # Group by BIN
    dplyr::group_by(BIN) %>% 
    # Calculate mean MQ within each bin
    dplyr::summarise(MEAN_MQ = mean(MQ, na.rm = T))
}) %>% 
  # bind into single data frame
  dplyr::bind_rows(.id = "CHR")

```

##### Final

```{r}
df_circos_500kb = cbind(dat_list$`500000`,
                        PI_WK = wk_list[["500000"]]$PI,
                        REPEAT_PROP = bin_intervals_500$REPEAT_PROP,
                        MEAN_MQ = mean_mq_500$MEAN_MQ)
# Add bin ends at the end of each chromosome to match real chromosome ends
df_circos_500kb$BIN_END = bin_intervals_500$BIN_END
```

```{r}
out_plot = here::here("docs/plots/nucleotide_diversity", "20211001_circos_500kb.png")
```

```{r}
png(out_plot,
    width = 20,
    height = 20,
    units = "cm",
    res = 500)

# Set parameters
## Decrease cell padding from default c(0.02, 1.00, 0.02, 1.00)
circos.par(cell.padding = c(0, 0, 0, 0),
           track.margin = c(0, 0),
           gap.degree = c(rep(1, nrow(med_chr_lens) - 1), 6))
# Initialize plot
circos.initializeWithIdeogram(med_chr_lens,
                              plotType = c("axis", "labels"),
                              major.by = 1e7,
                              axis.labels.cex = 0.25*par("cex"))

# Add MIKK Pi line
circos.genomicTrack(df_circos_500kb,
    panel.fun = function(region, value, ...){
  circos.genomicLines(region,
                      value[[2]],
                      col = "#49A379",
                      area = T,
                      border = karyoploteR::darker("#49A379"))
    },
    track.height = 0.1,
    bg.border = NA,
    ylim = c(0, 0.023))

circos.yaxis(side = "right",
           at = c(.01, .02),
           labels.cex = 0.25*par("cex"),
           tick.length = 2
           )

circos.text(0, 0.01,
        labels = expression(paste("MIKK (", pi, ")", sep = "")),
        sector.index = "1",
        facing = "clockwise",
        adj = c(.5, 0),
        cex = 0.4*par("cex"))  

# Add wild Kiyosu line
circos.genomicTrack(df_circos_500kb,
    panel.fun = function(region, value, ...){
  circos.genomicLines(region,
                      value[[4]],
                      col = "#7A306C",
                      area = T,
                      border = karyoploteR::darker("#8A4F7D"))
    },
    track.height = 0.1,
    bg.border = NA,
    ylim = c(0, 0.023))

circos.yaxis(side = "right",
           at = c(.01, .02),
           labels.cex = 0.25*par("cex"),
           tick.length = 2
           )

circos.text(0, 0.01,
        labels = expression(paste("Wild\nKiyosu (", pi, ")", sep = "")),
        sector.index = "1",
        facing = "clockwise",
        adj = c(.5, 0),
        cex = 0.4*par("cex"))  

# Add repeat content
circos.genomicTrack(df_circos_500kb,
  panel.fun = function(region, value, ...) {
    circos.genomicLines(region,
                        value[[5]],
                        type = "h",
                        col = "#0C1B33",
                        lwd = 0.4,
                        cex = 0.05,
                        baseline = 0)
  },
  track.height = 0.1,
  ylim = c(0,0.65),
  bg.border = NA)

circos.yaxis(side = "right",
           at = c(0, .5, 1),
           labels.cex = 0.25*par("cex"),
           tick.length = 5
           )

circos.text(0, 0.325,
        labels = "Repeat\ncontent",
        sector.index = "1",
        facing = "clockwise",
        adj = c(.5, 0),
        cex = 0.4*par("cex")) 

# Add mean mapping quality
circos.genomicTrack(df_circos_500kb,
  panel.fun = function(region, value, ...) {
    circos.genomicLines(region,
                        value[[6]],
                        type = "h",
                        col = "#FF6978",
                        lwd = 0.4,
                        cex = 0.05,
                        baseline = 20)
  },
  track.height = 0.07,
  ylim = c(20,65),
  bg.border = NA)

circos.yaxis(side = "right",
           at = c(20, 50),
           labels.cex = 0.25*par("cex"),
           tick.length = 5
           )

circos.text(0, 40,
        labels = "Mean MQ",
        sector.index = "1",
        facing = "clockwise",
        adj = c(.5, 0),
        cex = 0.4*par("cex")) 

# Print label in center
text(0, 0, "Nucleotide diversity in\nMIKK and wild Kiyosu medaka,\nwith repeat content\nand mapping quality\nin 500 kb windows")

circos.clear()

dev.off()
```

```{r}
knitr::include_graphics(out_plot)
```

# Per-individual

## Read in data

### MIKK

```{r}
# Set location of data
in_dir = "/nfs/research/birney/users/ian/mikk_genome/nucleotide_divergence/mikk/per_sample"

in_files = list.files(in_dir, full.names = T)
mikk_list = lapply(in_files, function(IN_FILE) {
  # Read in 
  readr::read_delim(IN_FILE,
                    delim = "\t",
                    col_types = c("ciiid")) %>% 
    # Remove MT
    dplyr::filter(CHROM != "MT") %>% 
    # Make CHR an integer
    dplyr::mutate(CHROM = as.integer(CHROM)) %>% 
    # Create middle of bin
    dplyr::mutate(BIN_MID = ((BIN_END - BIN_START - 1) / 2) + BIN_START )
})
names(mikk_list) = basename(in_files) %>%
  str_remove(".windowed.pi")

# reorder
ordered_lines = tibble::tibble("LINE" = names(mikk_list)) %>% 
  tidyr::separate(col = LINE, into = c("LINE", "SIB", "EXTRA"), sep = "_") %>% 
  dplyr::mutate(LINE = as.integer(LINE)) %>% 
  dplyr::arrange(LINE) %>% 
  tidyr::unite(col = "ORIGINAL", sep = "_", na.rm = T) %>% 
  dplyr::pull()

mikk_list = mikk_list[order(match(names(mikk_list), ordered_lines))]

# Bind rows
mikk_df = mikk_list %>% 
  dplyr::bind_rows(.id = "LINE") %>% 
  # factor LINE to order
  dplyr::mutate(LINE = factor(LINE, levels = names(mikk_list)))
```

### Wild

```{r}
# Set location of data
in_dir = "/nfs/research/birney/users/ian/mikk_genome/nucleotide_divergence/wild/per_sample"

in_files = list.files(in_dir, full.names = T)
wild_list = lapply(in_files, function(IN_FILE) {
  # Read in 
  readr::read_delim(IN_FILE,
                    delim = "\t",
                    col_types = c("ciiid")) %>% 
    # Remove MT
    dplyr::filter(CHROM != "MT") %>% 
    # Make CHR an integer
    dplyr::mutate(CHROM = as.integer(CHROM)) %>% 
    # Create middle of bin
    dplyr::mutate(BIN_MID = ((BIN_END - BIN_START - 1) / 2) + BIN_START )
})
names(wild_list) = basename(in_files) %>%
  str_remove(".windowed.pi")

# Bind rows
wild_df = wild_list %>% 
  dplyr::bind_rows(.id = "LINE")
```

### Combine

```{r}
pi_df = list("MIKK" = mikk_df,
             "WILD" = wild_df) %>% 
  dplyr::bind_rows(.id = "BREEDING")
```


## Plot

```{r}
mikk_list$`4_1` %>% 
  ggplot() +
    geom_histogram(aes(PI)) +
    facet_wrap(~CHROM)
```
### Chr1

```{r, fig.dim = c(8, 20)}
mikk_chr1 = mikk_df %>% 
  # Take only chromosome 1
  dplyr::filter(CHROM == 1) %>% 
  # Create Mb column
  dplyr::mutate(Mb = BIN_START / 1e6) %>% 
  ggplot() +
    geom_col(aes(Mb, PI, fill = PI)) +
    facet_grid(rows = vars(LINE), cols = vars(CHROM)) +
    theme_bw() +
    theme(text = element_text(size = 5)) +
    xlab("500 kb window start position (Mb)") +
    ylab(expression(paste("Nucleotide diversity (", pi, ")", sep = ""))) +
    ggtitle("MIKK nucleotide diversity (chr1)") +
    scale_fill_viridis(name = expression(pi))

mikk_chr1
```

```{r, eval = F}
ggsave(here::here("docs/plots/nucleotide_diversity/20211006_mikk_chr1_column.png"),
       plot = mikk_chr1,
       device = "png",
       width = 8,
       height = 20,
       units = "in",
       dpi = 400)
```

### All chromosomes

```{r}
purrr::map(1:24, function(CHR){
    # Create plot
    plot_df = mikk_df %>% 
      # Filter for target CHROM
      dplyr::filter(CHROM == CHR) %>% 
      # Create Mb column
      dplyr::mutate(Mb = (BIN_START - 1) / 1e6)
    out_plot = plot_df %>% 
      ggplot() +
        geom_col(aes(Mb, PI, fill = PI)) +
        facet_grid(rows = vars(LINE), cols = vars(CHROM)) +
        theme_bw() +
        theme(text = element_text(size = 5)) +
        ylim(c(0, max(mikk_df$PI))) +
        xlab("500 kb window start position (Mb)") +
        ylab(expression(paste("Nucleotide diversity (", pi, ")", sep = ""))) +
        ggtitle(paste("Chromosome ", CHR, sep = "")) +
        scale_fill_viridis(name = expression(pi), limits = c(0,max(mikk_df$PI)))
        
    
    # Adjust width dimensions of plot based on length of chromosome
    w_dim = mikk_df %>% 
      dplyr::filter(CHROM == CHR) %>% 
      dplyr::mutate(Mb = (BIN_START - 1) / 1e6) %>% 
      dplyr::pull(Mb) %>% 
      max(.)
    w_dim = w_dim * 0.26
    # Save
    ggsave(here::here("docs/plots/nucleotide_diversity/20211006_pi_per_chr", paste(CHR, ".png", sep = "")),
           plot = out_plot,
           device = "png",
           width = w_dim,
           height = 20,
           units = "in",
           dpi = 400)
})
```


```{r}
View(mikk_df %>% 
  dplyr::group_by(LINE) %>% 
  dplyr::summarise(mean(PI)))

View(wild_df %>% 
  dplyr::group_by(LINE) %>% 
  dplyr::summarise(mean(PI)))
```


